Sequence Specific Modeling of \emph{E.~coli} Cell-Free Protein Synthesis

### 2017-04-07 Paper_v1.tex
Working my way thru the results .... still ...

\subsubsection{Productivity}
The theoretical maximum mean productivity was inversely proportional to the protein length ($l_{P}$) and varied between 1 and 12 $\mu$M/h for the proteins sampled (Fig.~\ref{fig:Prof}A-B). The theoretical maximum productivity was within a standard deviation with and without
amino acid supplementation for each protein, but varied significantly between proteins.
This suggested the cell free reaction had sufficient metabolic precursors to power CFPS and synthesize each protein of interest with similar productivities.
However, the theoretical productivity was inversely proportional to protein size;
for instance, BMP10 (XX aa) had a productivity of about 2.5 $\mu$M/h whereas CAT (ZZ aa) had a productivity of about 12 $\mu$M/h.
To examine this further, the mean productivity was plotted against the carbon number of each protein (Fig.~\ref{fig:Prof}B).
The productivity and protein size were related by the power-law relationship $\alpha\times(l_{P})^{\beta}$, where $\alpha = 6.0\times 10^{6}$ and $\beta=-1.9$.
Interestingly, CAT did not obey the power-law relationship; the relatively high productivity of CAT was due to its T7 promoter.
CAT expressed under a P70a promoter followed the power-law correlation with a productivity of approximately 8.8 $\mu$M/h.
Taken together, the maximum productivity of a cell free reaction was found to be inversely proportional to protein size,
following a power-law relationship for proteins expressed under a P70a promoter.

@@

%To determine the carbon contribution from each substrate (glucose and amino acids) we examined the carbon flux going toward the production of deGFP for all three cases (Table \ref{tbl:yield_breakdown}).
%The cases supplemented with amino acids in the media showed the highest carbon yield.
%Finally, for the third case (without amino acids supplemented), the carbon yield was reduced to 51.1 $\pm$ 1.8\% for CAT, and the system used about twice the amount of glucose as in the first two cases.
%In this case, glucose was used to synthesize amino acids and provide the energy necessary to power transcription and translation; this trend was seen across all proteins.

\subsubsection{Yield}
The theoretical maximum mean carbon yield was inversely proportional to protein size and varied between 40-57\% for the proteins sampled (Fig.~\ref{fig:Prof}E-F).
The same inverse qualitative trend was observed as for energy efficiency.
There was a drop in carbon yield by about 7\% once amino acids were not available; this is most likely because glucose was utilized to synthesize the necessary amino acids for each protein as well as power the system.
For the first case, the system relied on a mixture of glucose and some amino acids for each protein with a carbon yield of 58.2 $\pm$ 2.3\% for CAT.
Once amino acid synthesis reactions were blocked in the network (second case), the carbon yield dropped to 57.1 $\pm$ 2.2\% for CAT.
Only the necessary amount of amino acids was used for the production of the protein of interest; thus, it may be hypothesized that all the glucose was used to power CFPS and did not contribute to the carbon yield.
In that case, the carbon yield without glucose contribution would be 100\%.
In the experimentally constrained case, CAT was produced with a carbon yield of 6.2\% compared to the theoretical maximum of 58.2\%.
This decrease in carbon yield suggests inefficiencies in CFPS that can potentially be improved.
ssFBA assumes a psuedo steady state; thus, intermediate metabolites cannot accumulate within the cell-free extract.
In addition, ssFBA is solved by maximizing the flux through the protein production reaction.
Therefore, carbon flux will travel through the network to optimize the maximum flux through the protein synthesis reaction.
In examining the experimental dataset, there is a high accumulation of organic acids, especially acetate, pyruvate and lactate.
The experimental performance could be improved by diverting this carbon toward the protein of interest by knockouts during the cell-free extract preparation.
% \begin{figure}[t!]
% \centering
% \includegraphics[width=1.00\textwidth]{./Figures/Yield.pdf}
% \caption{CFPS carbon yield of eight proteins for four cases. Amino acid uptake and synthesis (black), AA uptake without synthesis (dark grey), AA synthesis without uptake (light grey), and constrained by experimental measurements, for CAT only (white). A. Energy efficiency across the ensemble (error bars represent the 95\% CI). B. Mean energy efficiency versus the carbon number for each corresponding protein. Trendline of energy efficiency versus carbon number (black dotted line) for first two cases (y = -8.03$\cdot$10\textsuperscript{-3}x+63.83; R\textsuperscript{2} = 0.95) and trendline for AA synthesis without uptake (grey dotted line; y = -6.98$\cdot$10\textsuperscript{-3}x+58.86; R\textsuperscript{2} = 0.90). Dagger: constrained by experimental measurements and excluded from trendline.}
% \label{fig:Yield}
% \end{figure}
%Next we investigated the effect of the carbon number of each protein on the carbon yield (Fig.~\ref{fig:Yield}B).
%The proteins with the lowest carbon number had the highest yield and the higher carbon number proteins had a lower carbon yield within each case.
A single trendline was formulated for the cases where amino acids were supplied in the media since they had similar performance, while another trendline was formulated for when amino acids were not available (Fig.~\ref{fig:Prof}F).
The trendlines showed good predictability for estimating carbon yield for a certain range of protein size (R\textsuperscript{2} = 0.95 with amino acids and R\textsuperscript{2} = 0.90 without amino acids), regardless of the type of promoters used.
However, as the protein size increased, the carbon yield decreased, suggesting that larger proteins may be less feasible for cell-free production.
% Thus, we examined the parameters that had the most significant effect on cell-free productivity, energy efficiency, and carbon yield in order to optimize CFPS performance.

\subsection{Sensitivity analysis}
To better understand the effect of substrate utilization and the transcription/translation parameters on CFPS performance, we performed global sensitivity analysis on the productivity, energy efficiency and carbon yield for the CAT protein (Fig.~\ref{fig:SI}).
Surprisingly, RNAP and ribosome abundance had only a modest effect;
the translation elongation rate had the largest effect on protein productivity.
On the other hand, oxygen followed by glucose consumption had the largest influence on the energy efficiency and carbon yield, respectively.
The significance of transcription/translation parameters was robust to amino acid supplementation, with the translation rate being the most sensitive across all cases (Fig.~\ref{fig:SI}A).
This suggested that the translation elongation rate, and not transcriptional parameters, controlled productivity.
Underwood and coworkers showed that an increase in ribosome levels did not significantly increase protein yields or rates; however, adding elongation factors increased yields by 23\% at 30 minutes \cite{2005_underwood_biotech}.
In addition, Li et al. increased the productivity of firefly luciferase production by 5-fold by adjusting factors that affect transcription and translation such as elongation factors, ribosome recycling factor, release factors, chaperones, BSA, and tRNAs \cite{2014_li_PlosOne}.
In examining substrate utilization, glucose consumption was not important for productivity in the presence of amino acid supplementation.
However, its importance increased significantly when amino acids were not available.
On the other hand, amino acid consumption was only sensitive when amino acids synthesis reactions were blocked, as it was the only source of amino acids for CAT synthesis.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./figs/Fig-3-Sensitivity-Analysis.pdf}
\caption{Total order sensitivity of deGFP productivity (A) and energy efficiency (B) to specific uptake rates and transcription/translation parameters for three cases: amino acid uptake and synthesis (black), amino acid uptake without synthesis (dark grey), and amino acid synthesis without uptake (light gray). Error bars represent a 95\% CI on the sensitivity index.}
\label{fig:SI}
\end{figure}

The oxygen consumption rate was the most important factor controlling the energy efficiency of cell free protein synthesis (Fig.~\ref{fig:SI}B).
The transcription/translation parameters had the same trend as for productivity, where the translation rate was the most sensitive compared to the other transcription/translation parameters and showed significance across all cases.
Oxygen uptake was the most significant for energy efficiency, since it was responsible for oxidative phosphorylation, the most efficient pathway for energy generation.
Across all cases, substrate utilization (amino acid uptake and glucose) was shown to be the next most important as these substrates contributed to the energy efficiency of CAT synthesis.
In cases where amino acids were supplied, amino acid uptake was significant, as without it, energy was required to synthesize amino acids for the protein of interest.
Glucose uptake was only important when amino acids were not available, since it was the only source of carbon for protein synthesis and energy generation.
Jewett and coworkers have reported that oxidative phosphorylation still operated in cell-free systems, and that yield decreased from 1.5-fold to 4-fold when oxidative phosphorylation reactions were knocked out in pyruvate-powered CFPS \cite{Jewett:2008aa}.
It is unknown how active oxidative phosphorylation is compared to in \textit{in~vivo} systems.
To investigate this further we compared CAT carbon yield to oxidative phosphorylation flux (Fig.~\ref{fig:oxphos_yield}).
Interestingly, oxidative phosphorylation had a strong effect on the carbon yield in all cases.
The cases where amino acids were supplied followed the same trend: ranging from a carbon yield of 20\% to 58\%, depending on the oxidative phosphorylation activity.
Once amino acids were removed from the media, the carbon yield dropped to about 10\%, and reached a maximum of 52\%.
When amino acids are not avialable in the media, a lower carbon yield was expected for the same oxidative phosphorylation flux, since carbon must be utilized for energy generation and amino acid biosynthesis.
In all cases, whenever the carbon yield was below its theoretical maximum, there was an accumulation of acetate and lactate, resulting in the lower carbon yield.
The experimental dataset exhibits a mixture of acetate and lactate accumulation during CAT synthesis, which shows that CFPS is not operating in a fully aerobic state.
Glucose can not be fully oxidized and therefore fermentation pathways are used.
Oxidative phosphorylation relies on electron transport from membrane vesicles, however CFPS has no cell membrane, thus it is expected CFPS has limited oxidative phosphorylation activity.
The addition of phosphate showed an increase in CAT yield; however, it is unclear whether the addition of phosphate enhances oxidative phosphorylation, inhibits phosphatase reactions, or both \cite{Jewett:2008aa}.
Interestingly, the addition of NADH did not increase the rate of protein synthesis, since the concentration of ATP was most likely saturated.
An alternative strategy may be the inhibition of anaerobic processes in cell-free, in order to minimize unwanted byproducts such as acetate and lactate.
%Thus, it is interesting to see how much of the energetic needs of the system are met by oxidative phosphorylation and how much from anaerobic processes.
\begin{figure}[t!]
\includegraphics[width=0.7\textwidth]{./figs/Fig-4-OxyPhox-versus-Carbon-Yield.pdf}
\caption{deGFP carbon yield versus oxidative phosphorylation flux, across an ensemble of 1000 ssFBA solutions, for three cases: amino acid uptake and synthesis (black), amino acid uptake without synthesis (dark grey), and amino acid synthesis without uptake (light gray).}
\label{fig:oxphos_yield}
\end{figure}

@@

\subsubsection{Potential Alternative Metabolic Optima}
Despite the constraints by experimental measurements, it is difficult to calculate the physiological flux distribution of metabolism (Fig.~\ref{fig:norm})
For example, there is a high flux through the Entner-Douodoroff pathway, but this is likely non-physiological, and simply an artifact of the optimal solution of ssFBA.
A knockout of the Entner-Doudoroff pathway has no effect on the norm productivity of CAT compared to no knockouts (Fig.~\ref{fig:norm}A).
In addition, pairwise knockouts of Entner-Doudoroff and other subgroups (i.e. pentose phosphate pathway, cofactors, folate metabolism, etc.) in the network result in the same optimal solution of CAT productivity.
However, there is a difference in the flux distribution with these knockouts (Fig.~\ref{fig:norm}B); ssFBA will reroute the flux to optimize the objective function.
Interestingly, there are crucial subgroups that are detrimental to CAT productivity.
For example, a single group knockout of glycolysis/gluconoeogenesis or oxidative phosphorylation leads to no CAT production.
There were also cases of group knockouts that had no effect on productivity or its flux distribution such as isoleucine, leucine, and valine biosynthesis knocked out with histidine biosynthesis.
Flux balance analysis has been shown to have multiple alternative optimal solutions with flux variablitiy analysis and mixed-integer linear programming resulting in a poor depiction of the physiological flux distribution \cite{LEE2000711, Mahadevan2003264, Schuetz119}.
In our case, ssFBA reached the same optimal solution of CAT productivity for 39.9\% of the pairwise group knockouts, but had a different flux distribution than compared to no knockouts.
To determine which reactions occur in CFPS, adding thermodynamic feasibility constraints to reactions may result in a better depiction of the flux distribution \cite{Henry:2007,Hamilton:2013}.
It would also be interesting to track the carbon flux using C\textsuperscript{13} labeling in CFPS and constrain branch reactions in ssFBA to the resulting measurements, a method that has been shown to represent the flux distribution for \emph{in~vivo} processes well \cite{Zamboni:2009}.
\begin{figure}[t!]
\includegraphics[width=1.0\textwidth]{./figs/Fig-7-FluxDistribition-Analysis.pdf}
\caption{The norm effect of pairwise knockouts of subgroups in the cell-free network (A) CAT productivity norm compared to no knockouts in the experimentally constrained case. (B) Flux distribution norm compared to no knockouts in the experimentally constrained case.}
\label{fig:norm}
\end{figure}

>>>>

\subsubsection{Productivity}
The theoretical maximum mean productivity ($\mu$M/h) was inversely proportional to the protein length ($l_{P}$) and varied between 1 and 12 $\mu$M/h for the proteins sampled (Fig.~\ref{fig:Prof}A-B). The theoretical maximum productivity was within a standard deviation with and without
amino acid supplementation for each protein, but varied significantly between proteins.
For instance, BMP10 (XX aa) had a optimal productivity of approximately 2.5 $\mu$M/h whereas CAT (ZZ aa) had an optimal productivity of approximately 12 $\mu$M/h.
To examine the influence of protein length further, the mean optimal productivity was plotted against the carbon number of each protein (Fig.~\ref{fig:Prof}B).
The optimal productivity and protein length were related by the power-law relationship $\alpha\times(l_{P})^{\beta}$, where $\alpha = 6.0\times 10^{6}$ [UNITS?] and $\beta=-1.9$.
Interestingly, CAT did not obey the power-law relationship; the relatively high productivity of CAT was due to its T7 promoter.
However, CAT expressed under a P70a promoter followed the power-law correlation with a productivity of approximately 8.8 $\mu$M/h (predicted to be XX by the optimal productivity correlation). Taken together, the maximum optimal productivity of a cell free reaction was found to be inversely proportional to protein size,
following a power-law relationship for proteins expressed under a P70a promoter.

@@

%To determine the carbon contribution from each substrate (glucose and amino acids) we examined the carbon flux going toward the production of deGFP for all three cases (Table \ref{tbl:yield_breakdown}).
%The cases supplemented with amino acids in the media showed the highest carbon yield.
%Finally, for the third case (without amino acids supplemented), the carbon yield was reduced to 51.1 $\pm$ 1.8\% for CAT, and the system used about twice the amount of glucose as in the first two cases.
%In this case, glucose was used to synthesize amino acids and provide the energy necessary to power transcription and translation; this trend was seen across all proteins.
% W, the system relied on a mixture of glucose and some amino acids for each protein with a carbon yield of 58.2 $\pm$ 2.3\% for CAT.
% Once amino acid synthesis reactions were blocked in the network (second case), the carbon yield dropped to 57.1 $\pm$ 2.2\% for CAT.

\subsubsection{Carbon Yield}
The theoretical maximum mean carbon yield was inversely proportional to protein length and varied between 40\% and 57\% for the proteins sampled (Fig.~\ref{fig:Prof}E-F).
The relationship between the optimal carbon yield and protein length was linear; $m_{Y}\times(l_{P})+b_{Y}$ where $m_{Y} =  -8.03\times10^{-3}$ [UNITS] with supplementation, and $m_{Y} = -6.98\times10^{-3}$ [UNITS] without. The y-intercept $b_{Y} = 63.83$ [UNITS] with supplementation, and $b_{Y} = 58.86$ [UNITS] without.
The linear carbon yield models showed good predictability for estimating the optimal carbon yield for a range of protein lengths ($R^{2} = 0.95$ with amino acids
and $R^{2} = 0.90$ without supplementation), irrespective of the promoter used to control protein expression.
The effective yield models were approximately parallel, with a drop in carbon yield of approximately 7\% without supplementation.
Only the necessary amount of amino acids were used for the production of the protein of interest; thus, it may be hypothesized that all the glucose was used to power CFPS and did not contribute to the carbon yield.
In that case, the carbon yield without the glucose contribution would be 100\%.
However, in the experimentally constrained case, CAT was produced with a carbon yield of 6.2\% compared to the theoretical maximum of 58\%.
This decrease in carbon yield suggests inefficiencies in CFPS that can potentially be improved.
ssFBA assumes a psuedo steady state; thus, intermediate metabolites cannot accumulate within the cell-free extract.
In addition, ssFBA is solved by maximizing the flux through the protein production reaction.
Therefore, carbon flux will travel through the network to optimize the maximum flux through the protein synthesis reaction.
In examining the experimental dataset, there is a high accumulation of organic acids, especially acetate, pyruvate and lactate.
The experimental performance could be improved by diverting this carbon toward the protein of interest by knockouts during the cell-free extract preparation.

% \begin{figure}[t!]
% \centering
% \includegraphics[width=1.00\textwidth]{./Figures/Yield.pdf}
% \caption{CFPS carbon yield of eight proteins for four cases. Amino acid uptake and synthesis (black), AA uptake without synthesis (dark grey), AA synthesis without uptake (light grey), and constrained by experimental measurements, for CAT only (white). A. Energy efficiency across the ensemble (error bars represent the 95\% CI). B. Mean energy efficiency versus the carbon number for each corresponding protein. Trendline of energy efficiency versus carbon number (black dotted line) for first two cases (y = -8.03$\cdot$10\textsuperscript{-3}x+63.83; R\textsuperscript{2} = 0.95) and trendline for AA synthesis without uptake (grey dotted line; y = -6.98$\cdot$10\textsuperscript{-3}x+58.86; R\textsuperscript{2} = 0.90). Dagger: constrained by experimental measurements and excluded from trendline.}
% \label{fig:Yield}
% \end{figure}
%Next we investigated the effect of the carbon number of each protein on the carbon yield (Fig.~\ref{fig:Yield}B).
%The proteins with the lowest carbon number had the highest yield and the higher carbon number proteins had a lower carbon yield within each case.
% A single trendline was formulated for the cases where amino acids were supplied in the media since they had similar performance, while another trendline was formulated for when amino acids were not available (Fig.~\ref{fig:Prof}F).

% Thus, we examined the parameters that had the most significant effect on cell-free productivity, energy efficiency, and carbon yield in order to optimize CFPS performance.

\subsection{Global sensitivity analysis}
To better understand the effect of substrate utilization and the transcription/translation parameters on CFPS performance, we performed global sensitivity analysis on the productivity, energy efficiency and carbon yield for the CAT protein (Fig.~\ref{fig:SI}).
Surprisingly, RNAP and ribosome abundance had only a modest effect;
the translation elongation rate had the largest effect on protein productivity.
On the other hand, oxygen followed by glucose consumption had the largest influence on the energy efficiency and carbon yield, respectively.
The significance of transcription/translation parameters was robust to amino acid supplementation, with the translation rate being the most sensitive across all cases (Fig.~\ref{fig:SI}A).
This suggested that the translation elongation rate, and not transcriptional parameters, controlled productivity.
Underwood and coworkers showed that an increase in ribosome levels did not significantly increase protein yields or rates; however, adding elongation factors increased yields by 23\% at 30 minutes \cite{2005_underwood_biotech}.
In addition, Li et al. increased the productivity of firefly luciferase production by 5-fold by adjusting factors that affect transcription and translation such as elongation factors, ribosome recycling factor, release factors, chaperones, BSA, and tRNAs \cite{2014_li_PlosOne}.
In examining substrate utilization, glucose consumption was not important for productivity in the presence of amino acid supplementation.
However, its importance increased significantly when amino acids were not available.
On the other hand, amino acid consumption was only sensitive when amino acids synthesis reactions were blocked, as it was the only source of amino acids for CAT synthesis.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./figs/Fig-3-Sensitivity-Analysis.pdf}
\caption{Total order sensitivity of deGFP productivity (A) and energy efficiency (B) to specific uptake rates and transcription/translation parameters for three cases: amino acid uptake and synthesis (black), amino acid uptake without synthesis (dark grey), and amino acid synthesis without uptake (light gray). Error bars represent a 95\% CI on the sensitivity index.}
\label{fig:SI}
\end{figure}

% The transcription/translation parameters had the same trend as for productivity, where the translation rate was the most sensitive compared to the other transcription/translation parameters and showed significance across all cases.

The oxygen consumption rate was the most important factor controlling the energy efficiency of cell free protein synthesis (Fig.~\ref{fig:SI}B).
In addition to the oxygen consumption, amino acid utilization was also sensitive.
On the other hand, glucose consumption was only sensitive in the absence of amino acid supplementation,
since in the absence of amino acids, glucose acted as the sole source of carbon and energy for protein synthesis.
In the model, we assumed that ATP could be produced by both substrate level and oxidative phosphorylation.
Jewett and coworkers reported that oxidative phosphorylation still operated in cell free systems, and that the protein
yield decreased from 1.5-fold to 4-fold when oxidative phosphorylation reactions were inhibited in pyruvate-powered CFPS \cite{Jewett:2008aa}.
However, it is unknown how active oxidative phosphorylation is for a glucose powered cell free system.
Moreover, the connection between oxidative phosphorylation activity and other performance metrics, such as the carbon yield, is also unclear.

To investigate the connection between carbon yield and oxidative phosphorylation further,
we calculated the CAT carbon yield as a function of the oxidative phosphorylation flux (Fig.~\ref{fig:oxphos_yield}).
Oxidative phosphorylation had a strong effect on the carbon yield, both with and without amino acid supplementation.
In the presence of amino acid supplementation, the carbon yield ranged from 20\% to approximately 60\%,
depending upon the oxidative phosphorylation flux.
However, without amino acid supplementation, the carbon yield dropped to approximately 10\%, and reached a maximum of 50\%.
In the absence of supplementation, a lower carbon yield was expected for the same oxidative phosphorylation flux, as glucose was utilized for both energy generation and amino acid biosynthesis.
In all cases, whenever the carbon yield was below its theoretical maximum, there was an accumulation of both acetate and lactate.
The experimental dataset also exhibited a mixture of acetate and lactate accumulation during CAT synthesis, which suggested the CFPS reaction was not
operating with optimal oxidative phosphorylation activity.
Oxidative phosphorylation is a membrane associated process, however CFPS has no cell membrane.
Jewett et al hypothesized that membrane vesicles present in the CFPS reaction carry out oxidative phosphorylation [REF];
however, the number, size, protein loading and lifetime of these vesicles remains an open area of study.
Thus, we expect CFPS will have suboptimal oxidative phosphorylation capacity.

% The addition of phosphate showed an increase in CAT yield; however, it is unclear whether the addition of phosphate enhances oxidative phosphorylation, inhibits phosphatase reactions, or both \cite{Jewett:2008aa}.
% Interestingly, the addition of NADH did not increase the rate of protein synthesis, since the concentration of ATP was most likely saturated.
% An alternative strategy may be the inhibition of anaerobic processes in cell-free, in order to minimize unwanted byproducts such as acetate and lactate.
%Thus, it is interesting to see how much of the energetic needs of the system are met by oxidative phosphorylation and how much from anaerobic processes.
\begin{figure}[t!]
\includegraphics[width=0.7\textwidth]{./figs/Fig-4-OxyPhox-versus-Carbon-Yield.pdf}
\caption{deGFP carbon yield versus oxidative phosphorylation flux, across an ensemble of 1000 ssFBA solutions, for three cases: amino acid uptake and synthesis (black), amino acid uptake without synthesis (dark grey), and amino acid synthesis without uptake (light gray).}
\label{fig:oxphos_yield}
\end{figure}

@@

\subsubsection{Potential Alternative Metabolic Optima}
One of the interesting features of constraint based approaches is the ability to compute potentially alternative optimal solutions.
These solutions have the same objective value, e.g., the productivity but different underlying metabolic flux distributions.
Despite the constraints by experimental measurements, it is difficult to calculate the physiological flux distribution of metabolism (Fig.~\ref{fig:norm})
For example, there is a high flux through the Entner-Douodoroff pathway, but this is likely non-physiological, and simply an artifact of the optimal solution of ssFBA.
A knockout of the Entner-Doudoroff pathway has no effect on the norm productivity of CAT compared to no knockouts (Fig.~\ref{fig:norm}A).
In addition, pairwise knockouts of Entner-Doudoroff and other subgroups (i.e. pentose phosphate pathway, cofactors, folate metabolism, etc.) in the network result in the same optimal solution of CAT productivity.
However, there is a difference in the flux distribution with these knockouts (Fig.~\ref{fig:norm}B); ssFBA will reroute the flux to optimize the objective function.
Interestingly, there are crucial subgroups that are detrimental to CAT productivity.
For example, a single group knockout of glycolysis/gluconoeogenesis or oxidative phosphorylation leads to no CAT production.
There were also cases of group knockouts that had no effect on productivity or its flux distribution such as isoleucine, leucine, and valine biosynthesis knocked out with histidine biosynthesis.
Flux balance analysis has been shown to have multiple alternative optimal solutions with flux variablitiy analysis and mixed-integer linear programming resulting in a poor depiction of the physiological flux distribution \cite{LEE2000711, Mahadevan2003264, Schuetz119}.
In our case, ssFBA reached the same optimal solution of CAT productivity for 39.9\% of the pairwise group knockouts, but had a different flux distribution than compared to no knockouts.
To determine which reactions occur in CFPS, adding thermodynamic feasibility constraints to reactions may result in a better depiction of the flux distribution \cite{Henry:2007,Hamilton:2013}.
It would also be interesting to track the carbon flux using C\textsuperscript{13} labeling in CFPS and constrain branch reactions in ssFBA to the resulting measurements, a method that has been shown to represent the flux distribution for \emph{in~vivo} processes well \cite{Zamboni:2009}.
\begin{figure}[t!]
\includegraphics[width=1.0\textwidth]{./figs/Fig-7-FluxDistribition-Analysis.pdf}
\caption{The norm effect of pairwise knockouts of subgroups in the cell-free network (A) CAT productivity norm compared to no knockouts in the experimentally constrained case. (B) Flux distribution norm compared to no knockouts in the experimentally constrained case.}
\label{fig:norm}
\end{figure}
